---
title: "Exploratory Data Analysis (EDA) Exercises"
author: Mo Jmaileh
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
```


--- 

**Question 1**: 
Plot the distribution of price. Describe the distribution. Hint: You can also 
use `summary(diamonds$price)` to view a quantitative assessment of the distribution.  


---
```{r}
diamonds %>%
  ggplot(aes(x = price)) +
  geom_histogram()

```


**Question 2**:
Which of the 4Cs (carat, cut, clarity, color) is most important for predicting 
the price of a diamond?  

---
```{r}
# Of the 4C's, Carat is the most important predictor of price, as the median price continues to go up as carat increases. 
diamonds %>%
  ggplot(aes(x= cut, y = price)) +
  geom_boxplot() 

diamonds %>%
  ggplot(aes(x= carat, y = price)) +
   geom_boxplot(mapping = aes(group = cut_number(carat, 20)))

diamonds %>%
  ggplot(aes(x= clarity, y = price)) +
  geom_boxplot() 

diamonds %>%
  ggplot(aes(x= color, y = price)) +
  geom_boxplot() 

```

**Question 3**:
Visualize the distribution of carat, partitioned by price. Hint: Bin the price 
and show the carat distribution for each bin. You can bin the data with this command: 

`mutate(price_binned = cut(price, breaks=c(seq(0, max(price), 5000), Inf), dig.lab=10))`



``` {r}
diamonds %>%
  mutate(price_binned = cut(price, breaks=c(seq(0, max(price), 5000), Inf), dig.lab=10)) %>%
  ggplot(aes(x = price_binned, y = carat)) +
  geom_boxplot()
```
