---
title: "Data Wrangling Exercises"
auhtor: Cara Salyers
date: April 16, 2017
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, message = FALSE)
pdf(NULL)
```

The following six exercises are based on concepts covered in 
[Chapter 12][G&W 2017, 12] and [Chapter 13][G&W 2017, 13] of 
[R for Data Science][G&W 2017]. Use the `flights` and `who` datasets that come 
with the `nycflights13` and `tidyverse` packages to work on them, and start by 
loading the `tidyverse` and `nycflights13`.

```{r, echo = TRUE, eval = TRUE, warning = FALSE}
library(tidyverse)
library(nycflights13)
```

Also, run `?flights` and `?who` in the console to (re)familiarize yourself with
these datasets.

--- 

## Exercise 1
```{r, echo = TRUE, eval = TRUE, error = TRUE}
people <- tribble(
  ~name,            ~key,     ~value,
  #---------------/---------/--------
  "Ousmane Diaby",    "age",      23,
  "Ousmane Diaby", "height",     196,
  "Ousmane Diaby",    "age",      33,
  "Rich Eldh",        "age",      23,
  "Rich Eldh",     "height",     195
)

people %>% 
  spread(key, value)
```

Why can't you spread `people`? How could you add a new column to `people` to fix
the problem?

**Hint:** You can use `group_by()` and `row_number()` to create a `time` column
that uniquely identifies time periods in `people`. This `time` column will fix
the problem.

**Note:** This question is based on [12.3.3 #3][G&W 2017, 12.3.3]. It is
designed to help you hone your understanding of how 
[`spread()`][G&W 2017, 12.3.2] behaves with non-unique rows.


**ANSWER:**
You can't spread people because there are two ages listed for Ousmane. You can add a time column using row numbers to designate age at time 1 and time 2.
```{r Q1, echo = TRUE, eval = TRUE}
people %>% 
  group_by(name, key) %>% 
  mutate("time" = row_number()) %>% 
  spread(key, value)
```
---

## Exercise 2
```{r, echo = TRUE}
preg <- tribble(
  ~pregnant, ~male, ~female,
  #--------/------/---------
  "yes",        NA,      10,
  "no",         20,      12
)
```

Do you need to [gather][G&W 2017, 12.3.1] or [spread][G&W 2017, 12.3.2] `preg`
to tidy it? What variables are there in `preg`?

**Note:** This question is based on [12.3.3 #4][G&W 2017, 12.3.3].


**ANSWER:**
You do need to gather `preg` because it currently has `male` and `female` as variable names when there should be a vairable for `gender` instead and a variable for `count`
```{r Q2, echo = TRUE}
preg %>% 
  gather(`male`, `female`, key = "gender", value = "count")
```

---

## Exercise 3
```{r, echo = TRUE, eval = TRUE, warning = TRUE}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))
```

How would you explain this warning message in layman's terms to someone who
couldn't figure out what it means? Suppose they want to make sure that every
piece of the tibble makes its way into the result of the 
[`separate()`][G&W 2017, 12.4.1] function call. What could they do to remedy the 
situation?

**Note:** This question is based on [12.4.3 #1][G&W 2017, 12.4.3]. It is
designed to strengthen your ability to parse a column using 
[`separate()`][G&W 2017, 12.4.1].

**ANSWER:**
The error message indicates that there was an extra piece entered in the second set of letters because set 1 and 3 only had three letters, so the last one (g) was dropped. You can change the default setting for extra to equal merge so instead of dropping additional pieces they will be merged.
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = 'merge')
```

---

## Exercise 4
Both  [`separate()`][G&W 2017, 12.4.1]  and [`unite()`][G&W 2017, 12.4.2] have a
`remove` argument. What does it do? When does setting it to `FALSE` make sense?

**Note:** This question is based on [12.4.3 #2][G&W 2017, 12.4.3]. It is
designed to strengthen your ability to parse columns while making use of
arguments such as `remove`.


**ANSWER:**
The `remove` argument tells whether to remove the input column from the output data frame when separating or uniting. It makes sense to set this to false when you want to keep the original column because it would still be useful for analysis or informational purposes.

---

## Exercise 5
Using the [`who` dataset][G&W 2017, 12.6], calculate the total number of cases
of TB per year for China, India, and Bangladesh, then plot these statistics
over time. What country-year statistics, if any, surprise you?

**Hint:** Build on this code, which tidies `who`:

```{r, echo = TRUE}
who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = gsub("newrel", "new_rel", code)) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

**Note:** This question is based on [12.6.1 #4][G&W 2017, 12.6.1]. It is
designed to strengthen your ability to create a pipeline full of data wrangling
and visualization operations.


**ANSWER:**
The statistics for India in 2007 are a surprising spike that shoots up for one year, then drops back down, which could indicate something weird happening with the data. There is also a big jump from 2008 to 2009 for China.
```{r, echo = TRUE}
who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = gsub("newrel", "new_rel", code)) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1) %>% 
  group_by(country, year) %>%
  summarise(cases = sum(value)) %>%
  filter(country %in% c('China', 'India', 'Bangladesh')) %>% 
  ggplot(aes(x = year, y = cases, colour = country)) + 
  geom_point() + 
  geom_line()
```

---

## Exercise 6
Begin by following these steps:

1. Join observations in `flights` with those in `weather` by `origin`, `year`,
`month`, `day`, and `hour`.  
2. Bin `wind_gust` and calculate average `dep_delay` for each bin.  
3. Create a bar graph.

At what gust speeds are departure delays out of NYC the longest on average?

**Note:** This question is based on [13.4.6 #4][G&W 2017, 13.4.6]. It is
designed to strengthen your ability to merge and analyze relational data with
[`inner_join()`][G&W 2017, 13.4.2] and [`summarize()`][G&W 2017, 5.6].


**ANSWER:**
Departure delays are longest for when gust speeds of 35-40mph
```{r}
flights %>%
  inner_join(weather, by = c('origin', 'year', 'month', 'day', 'hour')) %>%
  mutate(wind_gust_binned = 
           cut(wind_gust, breaks = seq(0, max(wind_gust, na.rm = TRUE), 5))) %>%
  group_by(wind_gust_binned) %>%
  summarize(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = wind_gust_binned, y = dep_delay)) +
    geom_col()
  
```


[G&W 2017]: http://r4ds.had.co.nz/
[G&W 2017, 5.6]: http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise
[G&W 2017, 12]: http://r4ds.had.co.nz/tidy-data.html
[G&W 2017, 12.3.1]: http://r4ds.had.co.nz/tidy-data.html#gathering
[G&W 2017, 12.3.2]: http://r4ds.had.co.nz/tidy-data.html#spreading
[G&W 2017, 12.3.3]: http://r4ds.had.co.nz/tidy-data.html#exercises-22
[G&W 2017, 12.4.1]: http://r4ds.had.co.nz/tidy-data.html#separate
[G&W 2017, 12.4.2]: http://r4ds.had.co.nz/tidy-data.html#unite
[G&W 2017, 12.4.3]: http://r4ds.had.co.nz/tidy-data.html#exercises-23
[G&W 2017, 12.6]: http://r4ds.had.co.nz/tidy-data.html#case-study
[G&W 2017, 12.6.1]: http://r4ds.had.co.nz/tidy-data.html#exercises-25
[G&W 2017, 13]: http://r4ds.had.co.nz/relational-data.html
[G&W 2017, 13.4.2]: http://r4ds.had.co.nz/relational-data.html#inner-join
[G&W 2017, 13.4.6]: http://r4ds.had.co.nz/relational-data.html#exercises-28