---
title: "Data Wrangling Exercises"
author: "Zoe Zheng"
date: "April 17, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pdf(NULL)
```

The following 6 questions are based on concepts covered in Chapters 12-13 in R4DS. 
The questions will center around the `flights` dataset from the `nycflights13`, 
the `who` dataset from the `tidyr` package, and a few small example datasets. 
Start by loading the `tidyverse` and `nycflights13` packages. The `tidyr` package 
is loaded by default when the tidyverse is loaded.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

It is highly recommended to run `?flights` and `?who` from the R console to first 
familiarize yourself with the datasets.

--- 

**Question 1**:
Why does spreading this tibble fail? How could you add a new column to fix the problem? 
Hint: You can create an indicator column to uniquely identify each observation using 
the row number like this: `mutate(obs = row_number())`

```{r Q1, error=TRUE}
people <- tribble(
 ~ID, ~name,          ~key,    ~value,
#----|---------------|--------|------
 01, "Phillip Woods", "age",    45,
 01, "Phillip Woods", "height", 186,
 02, "Phillip Woods", "age",    50,
 03, "Jessica Cordero", "age",  37,
 03, "Jessica Cordero", "height", 156
)

people %>% 
  spread(key, value)
```

*This question is [12.3.1.#3](http://r4ds.had.co.nz/tidy-data.html#exercises-22) and grows your understanding of how `spread()` behaves with non-unique rows [(link)](http://r4ds.had.co.nz/tidy-data.html#spreading).*

---

**Question 2**:
Do you need to spread or gather the tibble below to tidy it? What are the variables?

```{r Q2}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

preg %>% 
  gather(`male`, `female`, 
         key = "gender", 
         value = "count")

preg

```

*This question is [12.3.1.#4](http://r4ds.had.co.nz/tidy-data.html#exercises-22) and grows your understanding of what "gathering" a dataset means [(link)](http://r4ds.had.co.nz/tidy-data.html#gathering).*

---

**Question 3**:
Running the following code generates a warning message. What does this message 
mean? Provide two ways to incorporate all of the data.

```{r Q3}
# tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
#   separate(x, c("one", "two", "three"))

#The warning message is "Too many values at 1 locations: 2", which is because 
#there are 4 four factors, d, e, f, g, in the second row, but there are only three 
#columns.

# method 1
tibble(x = c("a,b,c,NA", "d,e,f,g", "h,i,j,NA")) %>% 
  separate(x, c("one", "two", "three","four"))

#method 2
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra='merge')

```

*This question is [12.4.3.#1](http://r4ds.had.co.nz/tidy-data.html#exercises-23) and grows your ability to separate (parse) a column [(link)](http://r4ds.had.co.nz/tidy-data.html#separate).*

---

**Question 4**:
Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would 
you set it to FALSE?  
*This question is [12.4.3.#2](http://r4ds.had.co.nz/tidy-data.html#exercises-23) and grows your understanding of how arguments can be used for finer control when parsing a column [(link)](http://r4ds.had.co.nz/tidy-data.html#spreading-and-gathering).*

`remove` is used for removing columns when you unite or separate columns, and the
default setting is "TRUE". Setting it to FALSE makes sure you keep the original 
columns. 

---

**Question 5**:
Using the `who` dataset, find the total number of cases of TB each year for China, 
India, and Bangladesh, and plot the data over time. Do you notice anything surprising? 
Hint: First, run the code block below to generate a cleaned dataset and answer 
the question using that `clean_who` dataset.  

```{r Q5}
clean_who <- who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = gsub("newrel", "new_rel", code)) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)

total_tb <- clean_who %>% 
  group_by(country,year) %>% 
  summarise(total = sum(value)) %>% 
  filter(country %in% c("China", "India", "Bangladesh"))

ggplot(total_tb,aes(year,total,group = country)) + geom_line(aes(color = country))
```

There is a spike for China and India and a decine for Bangladesh at the same time.

*This question is based on [12.6.1.#4](http://r4ds.had.co.nz/tidy-data.html#exercises-25) and grows your ability to create a full wrangling pipeline and generate results [(link)](http://r4ds.had.co.nz/tidy-data.html#case-study).*

---

**Question 6**:
What weather conditions make it more likely to see a departure delay? Hint: Your 
answer should follow these steps:

```{r Q6}
flights_weather <- left_join(flights, weather, by = c("origin" = "origin", "year" = "year", "month" = "month", "day" = "day", "hour" = "hour"))

flights_weather %>% 
  mutate(wind_gust_binned = cut(wind_gust, breaks=seq(0, max(wind_gust,na.rm=T), 5)), 
         precip_binned = cut(precip, breaks=seq(0, max(precip,na.rm= T), .05))) %>%
  group_by(wind_gust_binned) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot(aes(wind_gust_binned, avg_delay)) + geom_col()
```

1. Join observations existing in both `flights` with `weather` based on `origin`, 
`year`, `month`, `day`, and `hour`
2. Bin `wind_gust` and find the average `dep_delay` for different levels of gusts
3. Create create a plot

It's most likely to see a departure delay when the wind gust is between 35 and 40.

*This question is based on [13.4.6.#4](http://r4ds.had.co.nz/relational-data.html#exercises-28) and grows your ability to use an inner join before summarizing data [(link)](http://r4ds.had.co.nz/relational-data.html#inner-join).*
