---
title: "Data Wrangling Exercises"
output: github_document
author: Paige Atkins 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following 6 questions are based on concepts covered in Chapters 12-13 in R4DS. 
The questions will center around the `flights` dataset from the `nycflights13`, 
the `who` dataset from the `tidyr` package, and a few small example datasets. 
Start by loading the `tidyverse` and `nycflights13` packages. The `tidyr` package 
is loaded by default when the tidyverse is loaded.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

It is highly recommended to run `?flights` and `?who` from the R console to first 
familiarize yourself with the datasets.

?flights
?who
--- 

**Question 1**:
Why does spreading this tibble fail? How could you add a new column to fix the problem? 
Hint: You can create an indicator column to uniquely identify each observation using 
the row number like this: `mutate(obs = row_number())`

```{r Q1, error=TRUE}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

#Spreading the tibble didnt not work as there were multiple values for one name, ie Phillip Woods is age 45 and age 50.

people %>% 
  group_by(name, key) %>% 
  mutate(id = row_number()) %>% 
  spread(key, value)

```





---

**Question 2**:
Do you need to spread or gather the tibble below to tidy it? What are the variables?

```{r Q2}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

#Yes, gather the tibble. Variables are pregnant and gender

preg %>% 
  gather(`male`, `female`, key = "gender", value = "count")

```

*This question is [12.3.1.#4](http://r4ds.had.co.nz/tidy-data.html#exercises-22) and grows your understanding of what "gathering" a dataset means [(link)](http://r4ds.had.co.nz/tidy-data.html#gathering).*

---

**Question 3**:
Running the following code generates a warning message. What does this message 
mean? Provide two ways to incorporate all of the data.

```{r Q3}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))


# "Expected 3 pieces. Additional pieces discarded in 1 rows [2]." = error message
# This means there is an additional data point in in row #2 and R deleted it. Fix this by:

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = 'merge')

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = 'drop', fill = 'right')
```

*This question is [12.4.3.#1](http://r4ds.had.co.nz/tidy-data.html#exercises-23) and grows your ability to separate (parse) a column [(link)](http://r4ds.had.co.nz/tidy-data.html#separate).*

---

**Question 4**:
Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would 
you set it to FALSE?  

# 'remove' determines if R should continue the default action of removing columns OR keep the original column of data points.


*This question is [12.4.3.#2](http://r4ds.had.co.nz/tidy-data.html#exercises-23) and grows your understanding of how arguments can be used for finer control when parsing a column [(link)](http://r4ds.had.co.nz/tidy-data.html#spreading-and-gathering).*
---

**Question 5**:
Using the `who` dataset, find the total number of cases of TB each year for China, 
India, and Bangladesh, and plot the data over time. Do you notice anything surprising? 
Hint: First, run the code block below to generate a cleaned dataset and answer the question using that `clean_who` dataset.  

```{r Q5}
clean_who <- who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = gsub("newrel", "new_rel", code)) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1) %>% 
  group_by(country, year) %>%
  summarise(cases = sum(value)) %>%
  filter(country %in% c('China', 'India', 'Bangladesh')) %>% 
  ggplot(aes(x = year, y = cases, colour = country)) + 
  geom_point() + 
  geom_line()

clean_who

#The most surprising thing is the hige spike in India in 2007

```




*This question is based on [12.6.1.#4](http://r4ds.had.co.nz/tidy-data.html#exercises-25) and grows your ability to create a full wrangling pipeline and generate results [(link)](http://r4ds.had.co.nz/tidy-data.html#case-study).*

---

**Question 6**:
What weather conditions make it more likely to see a departure delay? Hint: Your 
answer should follow these steps:

1. Join observations existing in both `flights` with `weather` based on `origin`, 
`year`, `month`, `day`, and `hour`
2. Bin `wind_gust` and find the average `dep_delay` for different levels of gusts
3. Create create a plot


flights %>%
  inner_join(weather, by = c('origin', 'year', 'month', 'day', 'hour')) %>%
  mutate(
    wind_gust_binned   = cut(wind_gust, 
                             breaks = seq(0, max(wind_gust, na.rm = TRUE), 5))
  ) %>%
  group_by(wind_gust_binned) %>%
  summarize(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = wind_gust_binned, y = dep_delay)) +
    geom_col()
