---
title: "Data Visualization Exercises"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following five questions are based on concepts covered in [Chapter 3 of R4DS][Grolemund and Wickham 2017, 3]. They can be answered using the `mpg` dataset from the `ggplot2` package. Start by loading the `tidyverse` package, which includes `ggplot2`.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

Run `?mpg` in the console to familiarize yourself with the dataset.

--- 

**Question 1**: Run `ggplot(data = mpg)`. What do you see?  

This question is [3.2.4#1][Grolemund and Wickham 2017, 3.2.4 Exercises]. It is designed to get you thinking about visualizations with [`ggplot()`][Grolemund and Wickham 2017, 3.2.2] from the ground (i.e., canvas) up.

---

**Question 2**: Make a scatterplot of `cyl` vs `displ`, then make a boxplot with the same variables. What additional information does the boxplot convey? 

**Hint:** Use `as.factor(cyl)` while specifying the boxplot.  

This question is based on [3.2.4#4][Grolemund and Wickham 2017, 3.2.4 Exercises]. It is designed to strengthen your ability to plot a continuous variable by a categorical variable.

---

**Question 3**: What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful? Create an alternative visualization that displays the data more effectively. 

**Hint:** Use `geom_bar()` with `position = "fill"`. `geom_bar()` is covered in [Section 3.8][Grolemund and Wickham 2017, 3.8].  

This question is based on [3.2.4#5][Grolemund and Wickham 2017, 3.2.4 Exercises]. It is designed to strengthen your ability to plot the joint distribution of two categorical variables.

---

**Question 4**: What geom would you use to draw:

 - A line chart? 
 - A boxplot?
 - A histogram?
 - An area chart? 
 
This question is [3.6.1#1][Grolemund and Wickham 2017, 3.6.1 Exercises]. It is designed to strengthen your knowledege of the various [`ggplot()` geom objects][Grolemund and Wickham 2017, 3.6].

---

**Question 5**: Will these two graphs look different? Why/why not?

```{r Q5, eval = FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

This question is [3.6.1#5][Grolemund and Wickham 2017, 3.6.1 Exercises]. It is designed to strengthen your understanding of inheritence with respect to the [`ggplot()`][Grolemund and Wickham 2017, 3.2.2] `mapping` argument.

[Grolemund and Wickham 2017, 3]: http://r4ds.had.co.nz/data-visualisation.html
[Grolemund and Wickham 2017, 3.2.2]: http://r4ds.had.co.nz/data-visualisation.html#creating-a-ggplot
[Grolemund and Wickham 2017, 3.2.4 Exercises]: http://r4ds.had.co.nz/data-visualisation.html#exercises
[Grolemund and Wickham 2017, 3.6]: http://r4ds.had.co.nz/data-visualisation.html#geometric-objects
[Grolemund and Wickham 2017, 3.6.1 Exercises]: http://r4ds.had.co.nz/data-visualisation.html#exercises-3
[Grolemund and Wickham 2017, 3.8]: http://r4ds.had.co.nz/data-visualisation.html#position-adjustments