---
title: "Exploratory Data Analysis (EDA) Exercises"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following three questions are based on concepts covered in 
[Chapter 7 of R4DS][G&W 2017, 7]. They can be answered using the `diamonds`
dataset from the `ggplot2` package. Start by loading the `tidyverse` package,
which includes `ggplot2`.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

Run `?diamonds` in the console to familiarize yourself with the dataset.

--- 

## Question 1
Explore the distribution of `price` with `geom_histogram()` and other functions.
Challenge yourself to go beyond the first plot you generate, which might might
look something like this:

```{r}
diamonds %>% 
  ggplot(aes(price)) + 
  geom_histogram()
```

Use prose to describe the distribution you utlimately discover. What are some
of the most surprising features of the distribution to you?

**Hint:** It might help if you first view a quantitative assessment of the
distribution with `summary(diamonds$price)`.

This question is based on [7.3.4#2][G&W 2017, 7.3.4]. It is designed to
strengthen your ability to illuminate the distribution of a continuous variable
with `geom_histogram()`.

---

## Question 2
Which of the 4Cs (i.e., `carat`, `cut`, `color`, and `clarity`) is most
important for predicting the `price` of a diamond?

This question is based on [7.5.1.1#1][G&W 2017, 7.5.1.1]. It is designed to
strengthen your ability to interpret variable relationships from plots (see
[here][G&W 2017, 7.5]).

---

## Question 3
Visualize the distribution of `carat`, partitioned by `price`. Then, after
revisiting [Faceting][G&W 2017, 3.5], use `facet_wrap()` to accomplish the same.

**Hint:** After binning `price`, plot the distribution of `carat` for each bin.
Binning `price` can be accomplished with the following command:

```{r eval = FALSE}
diamonds %>% 
  mutate(
    price_binned = cut(price, breaks = c(seq(0, max(price), 5000), Inf), 
                       dig.lab = 10)
  )
```

This question is based on [7.5.3#2][G&W 2017, 7.5.3]. It is designed to further
strengthen your ability to interpret variable relationships from plots (see
[here][G&W 2017, 7.5] again).

[G&W 2017, 3.5]: http://r4ds.had.co.nz/data-visualisation.html#facets
[G&W 2017, 7]: http://r4ds.had.co.nz/exploratory-data-analysis.html
[G&W 2017, 7.3.4]: http://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-13
[G&W 2017, 7.5]: http://r4ds.had.co.nz/exploratory-data-analysis.html#covariation
[G&W 2017, 7.5.1.1]: http://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-15
[G&W 2017, 7.5.3]: http://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-17