---
title: "Exploratory Data Analysis (EDA) Exercises"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following 3 questions are based on concepts covered in Chapters 7-8 in R4DS 
and can be answered using the `diamonds` dataset from the `ggplot2` package. Start by 
loading the `tidyverse` package, which includes `ggplot2`.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

It is highly recommended to run `?diamonds` from the R console to first familiarize 
yourself with the dataset.

--- 

**Question 1**: 
Plot the distribution of price. Describe the distribution. Hint: You can also 
use `summary(diamonds$price)` to get some hard numbers around the distribution.  
*This question is [7.3.4.#2](http://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-13) and grows your ability to plot a histogram to view a distribution [(link)](http://r4ds.had.co.nz/exploratory-data-analysis.html#visualising-distributions).*

---

**Question 2**:
Which of the 4Cs (carat, cut, clarity, color) is most important for predicting 
the price of a diamond?  
*This question is based on [7.5.1.1.#2](http://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-15) and grows your ability to interpret variable relationships from plots [(link)](http://r4ds.had.co.nz/exploratory-data-analysis.html#covariation).*

---

**Question 3**:
Visualize the distribution of carat, partitioned by price. Hint: Bin the price 
and show the carat distribution for each bin. You can bin the data with this command: 

`mutate(price_binned = cut(price, breaks=c(seq(0, max(price), 5000), Inf), dig.lab=10))`

In this question you are also encouraged to try and use faceting. Review Section 
[3.5](http://r4ds.had.co.nz/data-visualisation.html#facets) for details on facetting.
*This question is based on [7.5.3.#2](http://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-17) and grows your ability to interpret variable relationships from plots [(link)](http://r4ds.had.co.nz/exploratory-data-analysis.html#covariation).*
