---
title: "Data Transformation Solutions"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following questions are based on concepts covered in [Chapter 5][chapter 5]
of R4DS, and answers to them lie in the `flights` dataset of the nycflights13
package. Load the tidyverse and nycflights13 packages to start working on them.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(nycflights13)
```

--- 

**Question 1**:
Responding to this set of questions requires two steps: (a) finding flights that
meet certain criteria and (b) counting these subsets of flights.

 - How many flights flew into LAX?
 - How many flights flew out of LAX?
 - How many flights were longer than or equal to 2,000 miles in distance?
 - How many flights were destined for airports in the Los Angeles area (LAX, 
 ONT, SNA, PSP, SBD, BUR, or LGB), but did not originate out of JFK?

For each question, consider filtering the `flights` dataset, then counting the
number of rows that remain using `nrow()`. You might find it helpful to revisit
section [5.2.1][5.2.1] of R4DS, which is where Hadley and Garrett review
comparison operators. For the fourth question, consider using the `%in%`
operator as well as the exclamation mark (`!`), which you can read in your head
or vocalize as "not."

We developed this set of questions based on the first exercise in section
[5.2.4][5.2.4 Exercises] of R4DS. Each is designed to strengthen your ability to
use the dplyr verb [`filter()`][filter] effectively.

```{r include = TRUE, eval = TRUE}
# How many flights flew into LAX?
flights %>% 
  filter(dest == "LAX") %>% 
  nrow()

# How many flights flew out of LAX?
flights %>% 
  filter(origin == "LAX") %>% 
  nrow()

# This is a bit of a trick question. You can confirm that the correct answer is
# zero by running ?flights. The dataset description states: "On-time data for
# all flights that *departed* NYC (i.e. JFK, LGA or EWR)" (emphasis added). LAX
# is the airport code for Los Angeles International Airport, so zero flights in
# this dataset originated from LAX.

# How many flights were longer than or equal to 2,000 miles in distance?
flights %>% 
  filter(distance >= 2000) %>% 
  nrow()

# How many flights were destined for airports in the Los Angeles area (LAX, ONT, 
# SNA, PSP, SBD, BUR, or LGB), but did not originate out of JFK?
flights %>% 
  filter(
    dest %in% c("LAX", "ONT", "SNA", "PSP", "SBD", "BUR", "LGB"), 
    origin != "JFK"
  ) %>% 
  nrow()
```

---

**Question 2**:
How many flights were "ghost flights"? A "ghost flight" is defined as a flight 
that departed, but never arrived (i.e., has a missing value for `arr_time`).

We developed this question based on the third exercise in section [5.2.4][5.2.4
Exercises] of R4DS. It's designed to strengthen your understanding of how the
dplyr verb [`filter()`][filter] treats `NA` values.

```{r include = TRUE, eval = TRUE}
flights %>% 
  filter(!is.na(dep_time), is.na(arr_time)) %>% 
  nrow()
```

---

**Question 3**: 
How does [`arrange()`][arrange] treat missing values, and how could you sort all 
of the rows with a missing `arr_time` to the top of the dataset?

We developed this question based on the first exercise in section [5.3.1][5.3.1
Exercises] of R4DS. It's designed to strengthen your understanding of how the
dplyr verb [`arrange()`][arrange] treats `NA` values.

```{r include = TRUE, eval = TRUE}
# With the arrange() function, all non-missing values get sorted in ascending or
# descending fashion; then rows with missing values get displayed. The following
# code brings the missing values to the top with desc(is.na(arr_time)):

flights %>% 
  arrange(desc(is.na(arr_time)))

# Note: `is.na(arr_time)` converts `arr_time` into `TRUE/FALSE` values. When
# sorting `TRUE/FALSE` values in ascending order, `FALSE` gets listed first, so
# `is.na(arr_time)` needs to be sorted in descending order to lift the missing
# values to the top of the dataset.
```

---

**Question 4**:
What do you observe after running the code below? How does this behavior reflect 
how [`select()`][select] helpers deal with uppercase and lowercase matching by 
default? How can you override this default behavior? 

```{r include = TRUE, eval = FALSE}
select(flights, contains("TIME"))
```

We developed this question based on the fourth exercise in section [5.4.1][5.4.1
Exercises] of R4DS. It's designed to strengthen your ability to use the dplyr
verb [`select()`][select] effectively.

```{r include = TRUE, eval = TRUE}
# The code above selects columns in the flights dataset that contain the letter
# t followed by i, m, and e. Importantly, the variabe names could contain
# "TIME", "time", "TiMe", and so on, because the select() helper contains() is
# not case sensitive by default. The following code sets the `ignore.case`
# argument to `FALSE` to inhibit this functionality.

select(flights, contains("TIME", ignore.case = FALSE))
```

---

**Question 5**:
For each destination more than or equal to 2,000 miles away from NYC's airports,
compute the total number of minutes their arrivals were delayed. Then, determine
how much, as a percentage, each of these destinations contributed to the total
number of arrival-delay minutes that long-haul destinations (i.e., those more
than or equal to 2,000 miles away from NYC's airports) amassed in 2013. Which
three destinations top the list?

**Hint:** Consider using [`filter()`][filter] to find the flights that flew more
than or equal to 2,000 miles in distance and had positive `arr_delay` values.
Then, [`group_by()`][summarize] `dest`, use [`summarize()`][summarize] to
calculate the total number of arrival-delay minutes by destination, use
[`mutate()`][mutate] to convert this metric into a proportion, use
[`arrange()`][arrange] to lift the top destiations to the top of the list, and
use `head()` to select the top three.

We developed this question based on the fourth exercise in section [5.7.1][5.7.1
Exercises] of R4DS. It's designed to strengthen your ability to calculate
[per-group proportions of totals][grouped mutates].

```{r include = TRUE, eval = TRUE}
flights %>%
  filter(distance >= 2000, arr_delay > 0) %>%
  group_by(dest) %>%
  summarize(arr_delay_mins = sum(arr_delay)) %>%
  mutate(arr_delay_pct_of_total = arr_delay_mins / sum(arr_delay_mins)) %>%
  arrange(desc(arr_delay_pct_of_total)) %>% 
  head(3)
```

[5.2.1]: https://r4ds.had.co.nz/transform.html#comparisons
[5.2.4 Exercises]: http://r4ds.had.co.nz/transform.html#exercises-7
[5.3.1 Exercises]: http://r4ds.had.co.nz/transform.html#exercises-8
[5.4.1 Exercises]: http://r4ds.had.co.nz/transform.html#exercises-9
[5.7.1 Exercises]: http://r4ds.had.co.nz/transform.html#exercises-12
[chapter 5]: https://r4ds.had.co.nz/transform.html
[filter]: http://r4ds.had.co.nz/transform.html#filter-rows-with-filter
[arrange]: http://r4ds.had.co.nz/transform.html#arrange-rows-with-arrange
[select]: http://r4ds.had.co.nz/transform.html#select-columns-with-select
[mutate]: http://r4ds.had.co.nz/transform.html#add-new-variables-with-mutate
[grouped mutates]: http://r4ds.had.co.nz/transform.html#grouped-mutates-and-filters
[summarize]: http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise
