---
title: "Data Transformation Solutions"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, warning=FALSE, message=FALSE}
# required packages
library(tidyverse)
library(nycflights13)
```

--- 

**Question 1**:
The following questions ask you to find the number of flights that meet certain 
criteria. Hint: Count the number of rows using `nrow()` after you have filtered 
the dataset. A review of different comparison operators is presented in [5.2.2](http://r4ds.had.co.nz/transform.html#logical-operators).   
*This question is based on [5.2.4.#1](http://r4ds.had.co.nz/transform.html#exercises-7) and grows your ability to use [`filter()`](http://r4ds.had.co.nz/transform.html#filter-rows-with-filter).*

 - How many flights flew into `ORD`
 - How many flights flew out of `ORD`
 - How many flights are less than 100 miles?
 - How many flights were destined for Houston airports ('IAH' or 'HOU'), but did not 
 originate out of 'JFK'? Hint: Use the `%in%` operator. Also, use the 
 exclamation mark  (`!`), which means NOT.

**Answer:**

 - 17,283 flights - `flights %>% filter(dest == 'ORD') %>% nrow()`
 
 - 0 flights - `flights %>% filter(origin == 'ORD') %>% nrow()`
 
Note: This is a trick question. You can confirm the correct answer is zero by 
running `?flights`. The dataset description states: "On-time data for all flights 
that departed NYC (i.e. JFK, LGA or EWR)". 'ORD' is the airport code for Chicago 
O'Hare, so zero flights in this dataset originated from O'Hare.

 - 1,633 flights - `flights %>% filter(distance < 100) %>% nrow()`
 
 - 8,325 flights - `flights %>% filter(dest %in% c('IAH', 'HOU'), origin != 'JFK') %>% nrow()`

---

**Question 2**:
How many flights were "ghost flights"? A "ghost flight" is defined as a flight 
that departed, but has missing `arr_time` (never arrived?!).  
*This question is based on [5.2.4.#3](http://r4ds.had.co.nz/transform.html#exercises-7) and grows your knowledge of how [`filter()`](http://r4ds.had.co.nz/transform.html#filter-rows-with-filter) treats `NA` values.*

**Answer:**  
458 flights

```{r Q2}
flights %>% 
  filter(is.na(arr_time), !is.na(dep_time)) %>% 
  nrow()
```

---

**Question 3**: 
How are missing values treated by `arrange()`? How could you sort all rows with 
a missing `arr_time` to the top of the dataset?  
*This question is based on [5.3.1.#1](http://r4ds.had.co.nz/transform.html#exercises-8) and test how [`arrange()`](http://r4ds.had.co.nz/transform.html#arrange-rows-with-arrange) treats `NA` values.*

**Answer**:
With the `arrange()` function all non-missing values are sorted in ascending or 
descending fashion, and then the rows with missing values are displayed last. 

Here is the code to sort observations in descending order with the missing 
values at the top:

```{r Q3}
flights %>% 
  arrange(desc(is.na(arr_time)), desc(arr_time))
```

Note: `is.na(arr_time)` converts `arr_time` into `TRUE/FALSE` values. When sorting 
`TRUE/FALSE` values in ascending order the `FALSE`s are listed first, so we must sort 
those in descending order to make the missing rows where `is.na(arr_time)` is 
`TRUE` rise to the top of the dataset.

---

**Question 4**:
What do you observe after running the code below? How does this behavior reflect 
how select helpers deal with upper and lowercase matching by default? How can you 
change that default? 

```{r Q4, eval=FALSE}
select(flights, contains("TIME"))`
```

*This question is based on [5.4.1.#4](http://r4ds.had.co.nz/transform.html#exercises-9) and grows your ability to use [`select()`](http://r4ds.had.co.nz/transform.html#select-columns-with-select).*

**Answer**:
The current code fails to select any column names because the `contains()` function 
is case-sensitive by default. You can set the `ignore.case` argument to `TRUE` to 
find any match of "TIME" regardless of the case, like this:

```{r Q4a, eval=FALSE}
select(flights, contains("TIME", ignore.case = TRUE))
```

---

**Question 5**:
For each destination, compute the total minutes of delay and then determine what 
proportion of total delay minutes each destination represents. What are the top 3 
destinations by proportion of total and what proportions do each of the top 3 
destinations represent? Hint: Use `summarize()` and `arrange()` to quickly find 
the top 3. The `summarize()` command is reviewed in 
[5.6](http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise) and 
`arrange()` is reviewed in [5.3](http://r4ds.had.co.nz/transform.html#arrange-rows-with-arrange).  
*This question is based on [5.7.1.#4](http://r4ds.had.co.nz/transform.html#exercises-12) and grows your ability to calculate a per group percentage of total [(link)](http://r4ds.had.co.nz/transform.html#grouped-mutates-and-filters).*

**Answer**:
 
 1. ORD = 5.4%
 2. ATL = 5.0%
 3. SFO = 3.9%

```{r Q5}
flights %>%
  filter(dep_delay > 0) %>%
  group_by(dest) %>%
  summarize(delay_mins = sum(dep_delay)) %>%
  mutate(delay_pct_of_total = delay_mins / sum(delay_mins)) %>%
  arrange(-delay_pct_of_total)
```
